Models

### 2022-12-07 analysis/05_model_metrics.Rmd
Model studies individually

Function approach to model
```{r}
get_model <- function(metric_name, formula, model_id, data, ...){
  print(c(metric_name, model_id))
  qlmer <- quietly(lmer)

  fit_data <- data %>%
    filter(metric == metric_name) %>%
    mutate(year_scaled = scale(year),
           duration_scaled = scale(duration, scale = FALSE),
           startyear_scaled = scale(startyear, scale = FALSE),
           taxa = as.factor(taxa),
           realm = as.factor(realm),
           climate = as.factor(climate))

@@

@@

  #get study level conditional estimates and add to over all estimates
  study_ests <- broom.mixed::tidy(model, effects="ran_vals") %>%
    filter(group == filter_name, term == "year_scaled") %>% 
    select(-c(effect, group, term)) %>%
    rename(cond.std.error = std.error, cond.estimate = estimate) %>%
    mutate(estimate = cond.estimate + year_scaled_est,
      cond.var = cond.std.error*cond.std.error, 
           var = cond.var + year_scaled_var,
           std.error = sqrt(var),
           upr.ci = estimate + (2*std.error),
           lwr.ci = estimate - (2*std.error),
           sig = case_when(
             lwr.ci < 0 & upr.ci > 0 ~ FALSE,
             TRUE ~ TRUE
           ),
      metric = metric,
      model_id = model_id)
}

@@

# study_slopes_wide <- study_slopes %>%
#   select(study_id, metric, estimate) %>%
#   pivot_wider(names_from = "metric", values_from = "estimate")
```

@@

>>>>

Function approach to model
```{r}
get_model <- function(metric_name, formula, model_id, data, study_id, ...){
  print(c(metric_name, model_id))
  qlmer <- quietly(lmer)

  if (missing(study_id)){
    fit_data <- data %>%
      filter(metric == metric_name) %>%
      mutate(year_scaled = scale(year),
             duration_scaled = scale(duration, scale = FALSE),
             startyear_scaled = scale(startyear, scale = FALSE),
             taxa = as.factor(taxa),
             realm = as.factor(realm),
             climate = as.factor(climate))
  } else{
    fit_data <- data %>%
      filter(metric == metric_name) %>%
      mutate(year_scaled = scale(year),
             duration_scaled = scale(duration, scale = FALSE),
             startyear_scaled = scale(startyear, scale = FALSE),
             taxa = as.factor(taxa),
             realm = as.factor(realm),
             climate = as.factor(climate)) %>%
      filter(study_id == study_id)
  }

@@

# get number of observations for each model
# lmm_obs <- model_table %>% 
#   filter(model_id == "base") %>% 
#   select(metric, model) %>% 
#   rowwise() %>% 
#   mutate(nobs = nobs(model)) %>% 
#   select(-model) %>% 
#   distinct()

@@

  #get study level conditional estimates and add to over all estimates
  study_ests <- broom.mixed::tidy(model, effects="ran_vals") %>%
    filter(group == filter_name, term == "year_scaled") %>% 
    select(-c(effect, group, term)) %>%
    rename(cond.std.error = std.error, cond.estimate = estimate) %>%
    mutate(estimate = cond.estimate + year_scaled_est,
      cond.var = cond.std.error*cond.std.error, 
           var = cond.var + year_scaled_var,
           std.error = sqrt(var),
           upr.ci = estimate + (1.96*std.error),
           lwr.ci = estimate - (1.96*std.error),
           sig = case_when(
             lwr.ci < 0 & upr.ci > 0 ~ FALSE,
             TRUE ~ TRUE
           ),
      metric = metric,
      model_id = model_id)
}

@@

usethis::use_data(study_slopes)
```

@@

#### Looking at BBS specifically to compare slopes and significance
```{r}
bbs_models <- map_dfr(metric_names[metric_names != "S"], 
                      ~get_model(data = model_data, 
                                 metric_name = .x, "value ~ year_scaled + (year_scaled|rarefyID)", 
                                 model_id = "base",
                                 study_id = 195))

bbs_models %>% 
  filter(term == "year_scaled") %>% 
  rowwise() %>% 
  mutate(nobs = nobs(model)) %>% 
  select(-model, -model_id, -formula, -n_ts, -warning, -group) %>% 
  relocate(metric) %>% 
  arrange(metric) %>% 
  write_csv(here::here("data/bbs_slopes_indvmodels.csv"))
```
